---
title: "RStudio Data Modeling Review"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, cache = F}
knitr::opts_chunk$set(
  echo = TRUE,
  error = TRUE,
  warning= FALSE,
  message= FALSE)
```

# Getting started
```{r}
library(tidyverse) # just run this code (no points)
library(tidymodels) # just run this code (no points)
library(corrplot)# just run this code (no points)
?nycflights13::flights# run this code to learn more about the dataset we will use in this review (no points)
flights_mod <- read_csv("flights_mod_review.csv") |> 
  mutate(year= as_factor(year), month=as_factor(month), day=as_factor(day))# just run this code (no points). You must run this code after running library(tidymodels and tidyverse) in order to replicate the results.
glimpse(flights_mod)# just run this code to explore the data more(no points).
```

1. Using the flights_mod dataset, write the code to create a correlation matrix using `arr_delay`, `dep_delay`, `distance`, `air_time`, `gain`  and `speed`. Make sure to name the matrix as flights_corr_matrix and to use only "complete.obs".
```{r}

flights_corr_matrix <- flights_mod |> 
  select(arr_delay, dep_delay, distance, air_time, gain  , speed) |> 
  cor(use ="complete.obs" )
```


2. Visualize the correlation matrix with a correlation chart.
```{r}
corrplot(flights_corr_matrix)
```


3. Write the code to set a seed for the data splitting task. Then create an initial split that allocate 80% of the data to the training set and 20% to the test set. Finally, using that split create a `flights_train` and a `flights_test` dataset.
```{r}
set.seed(123) 
flights_split <- initial_split(flights_mod, prop=0.8)
flights_train <- training(flights_split)
flights_test <- testing(flights_split)
```


4. Use the flights_train dataset to complete the following:

a. Create a new recipe and name the recipe as flights_recipe1.
  - Make sure that arr_delay is your dependent variable and that you use the `dep_delay`, `distance`, `air_time`, `origin`, `carrier`, `year`, `month`, `day` variables a predictors.
  - Impute missing values for the `dep_delay` using the average.
  - Log transform (base 10) all the numeric independent variables 
  - Create dummy variables for all the nominal predictors in the model. 
```{r}
flights_recipe1 <- recipe(arr_delay~dep_delay+distance+air_time+origin+ carrier+year+month+day, data = flights_mod) |> 
  step_impute_mean(dep_delay) |> 
  step_log(all_numeric_predictors(), base=10) |> 
  step_dummy(all_nominal_predictors())
```

b. Create a new recipe and name the recipe as flights_recipe2.
  - Make sure that arr_delay is your dependent variable and that you use the `dep_delay`, `distance`, `air_time`, `gain`, `speed` variables as predictors.
  - Standardize all the numeric independent variables.
  - Create an interaction term between  `distance` and `air_time`. 
```{r}

flights_recipe2 <- recipe(arr_delay~dep_delay+distance+air_time+gain+speed, data = flights_mod) |> 
  step_normalize(all_numeric_predictors()) |> 
  step_interact( ~ distance:air_time)
```


5.	Complete the following:
a.	Specify a linear regression model named "mod_linear_reg" with "lm" engine and with "regression" as mode.
```{r}
mod_linear_reg <- linear_reg() |> 
  set_engine("lm") |> 
  set_mode("regression")
```

b.	Specify a ridge regression model named "mod_linear_ridge" with "glmnet" as engine and with "regression" as mode.
```{r}
mod_linear_ridge <- linear_reg(penalty = 0.1, mixture=0) |> 
  set_engine("glmnet") |> 
  set_mode("regression")
```

6.	Complete the following:
a. Define	"reg_workflow_recipe1", make sure this workflow uses flights_recipe1 and mod_linear_reg model. Make sure to also fit the workflow without creating additional objects.
```{r}
reg_workflow_recipe1 <- workflow() |> 
  add_recipe(flights_recipe1) |> 
  add_model(mod_linear_reg)

fit( reg_workflow_recipe1, data = flights_train)
```


b. Define	"ridge_workflow_recipe1", make sure this workflow uses flights_recipe1 and mod_linear_ridge model. Make sure to also fit the workflow without creating additional objects.
```{r}
ridge_workflow_recipe1 <- workflow() |> 
  add_recipe(flights_recipe1) |> 
  add_model(mod_linear_ridge)

fit( ridge_workflow_recipe1, data = flights_train)

```


7. Complete the following:

a) Make sure to make predictions using reg_workflow_recipe1. Assign the predictions to an object named recipe1_reg_predictions. Finally make sure that the arrival delay predictions are added to the original flights_test dataset.

```{r}
recipe1_reg_predictions <- predict(reg_workflow_recipe1, new_data = flights_test) |> 
  bind_cols(flights_test)

```

b) Make sure to make predictions using ridge_workflow_recipe1 Assign the predictions to an object named recipe1_ridge_predictions. Finally make sure that the arrival delay predictions are added to the original flights_test dataset.

```{r}

```

8. Complete the following:

a. Compute and extract RMSE, MAE and RSquared associated with recipe1_reg_predictions. Make sure you can identify that the metrics belong to your `recipe1_reg_model`.

```{r}

```

b. Compute and extract RMSE, MAE and RSquared associated with recipe1_ridge_predictions. Make sure you can identify that the metrics belong to your `recipe1_ridge_model`.

```{r}

```

c. Which one is the best model? Please justify your choice.


9. Complete the following:

a. Write the code to show all the results of the reg_workflow_recipe1 model in a tidy format.

```{r}

```

b. Write the code to show all the results of the ridge_workflow_recipe1 model in a tidy format.

```{r}

```


10 [Bonus]. Provide a complete interpretation of the results of the reg_workflow_recipe1 model. Complete interpretation includes interpretation of coefficients, standard error, statistic and p-value. You should just focus on the `month` of July and `origin` variables.



## Submitting Your Code

To submit your code and your answers, simply copy and paste only the code needed to solve the tasks below each question available at this form (https://forms.office.com/r/1Er4iUPmgt). If you want answer the open ended questions (if any) directly in the form but make sure to don't leave them empty. 



