---
title: "Activity -2"
date: "`r format(Sys.time(), '%d %B, %Y')`" 
output: 
  html_document:
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(nycflights13)
knitr::opts_chunk$set(echo = FALSE)
```

# A Semester Project made by Custom Designed Activities


We continue working on the Dr. B & Class  project. During the entire semester you will apply what we learned in class in a set of tasks and scenarios custom designed for you. Please remember that it is important that the code that you submit is your own code and not somebody else work. It is fine to make mistakes but only by practicing in RStudio you can get a better grasp of the software. I also want you to try building your document as an official report for a potential company (Dreaming Diamonds LLC) for which you are getting to know and explore the diamonds dataset (e.g., spend time on storytelling, commenting results and providing insights and conclusions when possible). 
 
Complete all the activities starting at the bottom of this document (below the placeholder). Finally, once you have finished them, delete all the irrelevant content from this document and the instructions and make sure to rename this file as “week3_activities_your_name” (e.g., week3_activities_biagio_palese). Once you are done submit it in the Week3_activities folder in the Week 3 module on Blackboard. 

**Summary of this week tasks:**

This week we learn how to combine data manipulations using pipes. In an ideal world, you would receive a dataset ready for analysis. However, this is almost never the case. Cleaning and wrangling data are critical skills of data scientists and should be performed before moving to modeling. So, let’s practice what we covered this week:
- Advanced manipulations using pipes


## Week Minimum Requirements
**You must use the pipe (|>) to complete each task and do not create any intermediate object.**

1.	Load the tidyverse package.
```{r}
library(tidyverse)

```

2.	Use the diamonds dataset and complete the following:
a.	Group the dataset using the cut variable.
b.	Compute the following descriptive statistics for the carat variable: minimum, average, standard deviation, median, maximum.
c.	Produce the count of how many diamonds have each cut.
d.	What is the cut with the lowest number of observations in the dataset? Ans:fair 
What is the cut with the largest number of observations in the dataset? Ans:Ideal 
What is the cut with the highest average carat?ANS: Fair 
What is interesting about this analysis?
Ans:The data shows that the most popular cut, called "Ideal," has the highest number of diamonds (21551), indicating its popularity among consumers. Though Ideal cut diamonds are the most common, they tend to have a lower average carat weight compared to other cuts like Premium and Fair. This suggests that consumers may prioritize the quality of the cut over the size of the diamond, preferring smaller diamonds with superior cuts rather than larger ones with potentially inferior cuts.

```{r}
diamonds |> 
  group_by(cut) |> 
  summarise(minimum_carat= min(carat,na.rm = T),
            average_carat=mean(carat,na.rm = T), 
            standard_deviation_carat=sd(carat,na.rm = T),
            median_carat= median(carat,na.rm = T),
            maximum_carat=max(carat,na.rm = T),
            no_of_diamonds=n()) |> 
  arrange(no_of_diamonds)
```

3.	Use the diamonds dataset (?diamonds to familiarize again with it) and complete the following:
a.	Keep in the diamonds dataset only the carat, cut and price columns.
b.	Sort the dataset from the highest to the lowest price.
c.	Compute a new column named “price_per_carat” and equal to price/carat.
d.	Keep in the diamonds dataframe only the observations with price_per_carat above 10000$ and with a Fair cut. 
e.	How many observations are left in the dataset?Answer:5
What is the highest price per carat for a diamond with fair cut? Answer: 10909 
What is interesting about this analysis? 
ANSWER: One intriguing discovery is that diamonds with the same carat weight can have significantly different prices per carat. This suggests that factors beyond just carat weight, such as color, clarity, or dimensions, may heavily influence diamond pricing. Further analysis exploring these additional factors could provide valuable insights into why certain diamonds command higher prices per carat despite having identical carat weights.
```{r}
?diamonds
diamonds |> 
  select(carat, cut , price) |> 
  arrange(desc(price)) |> 
  mutate(price_per_carat=price/carat) |> 
  filter(price_per_carat>10000 & cut=="Fair") |> 
  arrange(desc(price_per_carat))
```

4.	Use the diamonds dataset and complete the following:
a.	Group the dataset using the color variable.
b.	Compute the following descriptive statistics for the price variable: minimum, average, standard deviation, median, maximum.
c.	Produce the count of how many diamonds have each color.
d.	Sort the data from the highest median price to the lowest. 
e.	What is the color with the lowest number of observations in the dataset? ANS: J 
What is the color with the largest number of observations in the dataset? ANS: G 
What is the color with the highest median price? ANS: J 
What is interesting about this analysis?
ANSWER: Despite common perceptions that certain colors might command consistently higher prices, the median price for J color diamonds is unexpectedly higher compared to other colors like D and E, which are traditionally considered higher quality due to their near-colorlessness. This finding suggests that factors beyond color, such as clarity, cut, or other characteristics, may play a significant role in determining diamond prices. Further investigation into these factors could provide valuable insights into consumer preferences and market dynamics within the diamond industry.
```{r}
diamonds |> 
  group_by(color) |> 
  summarise(minimum_price=min(price,na.rm = T),
            average_price=mean(price,na.rm = T), 
            standard_deviation_price=sd(price,na.rm = T), 
            median_price=median(price,na.rm = T), 
            maximum_price=max(price,na.rm = T), 
            no_of_diamonds=n()) |> 
  arrange(desc(median_price)) 
  
  
```

5.	Use the diamonds dataset and complete the following:
a.	Keep in the diamonds dataset only the clarity, price, x, y and z columns.
b.	Compute a new column named “size” and equal to x*y*z.
c.	Compute a new column named “price_by_size” and equal to price/size.
d.	Sort the data from the smallest to the largest price_by_size. 
e.	Group the observations by clarity.
f.	Compute the median price_by_size per each clarity.
g.	Keep in the dataset only observations with clarity equal to “IF” or “I1”.
h.	What is the median price_by_size for diamonds with IF clarity? ANS:19.0
What is the median price_by_size for diamonds with I1 clarity?ANS: 17.9 
Does is make sense that the median price_by_size for the IF clarity is bigger than the one for the I1 clarity? Why? 
ANS: The higher median price_by_size for IF clarity diamonds compared to I1 clarity diamonds is consistent with their differences in clarity grade, rarity, market demand, and perceived value.
```{r}
diamonds |> 
  select( clarity, price, x, y , z) |> 
  mutate(size=x*y*z, price_by_size=price/size) |>
  arrange(price_by_size) |> 
  group_by(clarity) |> 
  summarise(median_price_by_size = median(price_by_size)) |> 
  filter(clarity=="IF"|clarity=="I1")
  
  
```

## Submit your work
Test your code in RStudio and verify it solve the activities correctly before submitting using this link: https://forms.office.com/r/A3fxeDKmA3. Please make sure to copy and paste code and response directly in the form. We don't need you to upload your RMD file anymore.

